Probability Aggregation: Dynamic Hierarchical Modeling of Expert Beliefs in Geopolitical Events

### 2013-08-21 paper.tex
Edit
Added Barb and Phil. Also some cleanup.

\begin{aug}
\author{\fnms{Ville A.} \snm{Satop\"a\"a}\corref{}\ead[label=e1]{satopaa@wharton.upenn.edu}},
\author{\fnms{Shane T.} \snm{Jensen}\ead[label=e3]{stjensen@wharton.upenn.edu}}
\and
\author{\fnms{Lyle H.} \snm{Ungar}\ead[label=e2]{ungar@cis.upenn.edu}}

@@

 \affiliation{Department of Statistics\\ The Wharton School of the University of Pennsylvania   \printead{e1,e3}}

\affiliation{Department of Computer and Information Science\\ University of Pennsylvania, \printead{e2}}
 \address{Philadelphia, PA 19104- 6340, USA\\  \printead{e1}}

@@

\section{Introduction}
\label{intro}
Individual experts can differ radically from one another in their abilities to assess probabilities of future events. The quality of their assessments is typically evaluated with an important attribute called \textit{calibration}. An expert is said to be well-calibrated if 
the proportion of events that actually occurred is, say, 60\% for all those events for which he assessed a probability of 0.60. 
%the long-run frequency of occurrence of the events agrees with his probability forecasts. This means that, for instance, among all those events for which an expert assesses a probability of  0.60, the proportion of events that actually occur is 60\%.  
Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), some personal differences among the experts have been found to associate with calibration. In particular, \citet{wright1994coherence} argue that a higher self-rated expertise on the subject matter implies better calibration.

%Therefore self-reported expertise should be taken in to account when modeling probability forecasts given by a diverse group of experts. 
Unfortunately, calibration by itself does not guarantee usefulness of a probability assessment. For instance, if the empirical frequency of  the last 10 years of a relatively stationary process is 45\%, the expert can assess a constant probability of 0.45 and turn out well-calibrated. His probability, however, is not very useful in decision making. In general, a probability assessments that is near 0.0 or 1.0, is considered more informative, and an expert who makes such assessments is called \textit{sharp}.
%called \textit{smooth} or \textit{not sharp}. If, on other hand, most of his probabilities are near 0.0 or 1.0, he is deemed \textit{sharp}. 
%This notion leads to a second important attribute called \textit{sharpness}, which is a statement about the distribution of the assessed probabilities. If most of the probabilities are near 0.0 and 1.0, the assessor is said to be sharp.
In the forecasting literature sharpness is often combined with calibration into a single goal of maximizing sharpness subject to calibration (see, e.g., \citet{raftery2005using, Ranjan08}). An expert that meets this goal is considered to give perfect probability forecasts (\citet{gneiting2004verifying}). Even though expecting perfect forecasts from an expert is not reasonable in practice, this criterion has led to a wide range of novel and insightful observations in probability forecasting.

@@

Therefore many complex areas of probability aggregation still remain rather unexplored. 
%Therefore this paper makes an advance towards rather unexplored areas of probability aggregation by modeling forecasting 
%This paper introduces an aggregation procedure that allows multiple probability assessments from a single expert before the event resolves. 
%The context considered in this paper, on the other hand, allows the experts to update their probability assessments over a period of time before the future event resolves.
%
%considering a dynamic context with a time period during which the experts are allowed to update their beliefs. 
For instance, consider an investor aiming to assess whether a stock index will finish trading at or above some threshold on a given date. In order to maximize his overall predictive accuracy, he may consult a group of experts every day and adjust his estimate of the aggregate probability accordingly. Since the experts are allowed to update their probability assessments on a daily basis, the aggregation should be performed by taking into account the correlation and time-series nature of their advice. Analyzing this kind of future events under daily updating of the forecasts is the main focus of this paper. The core research goal is to construct an interpretable time-series model that incorporates self-reported expertise in order to estimate a sharp and well-calibrated version of the crowd belief. The model is then used for
\begin{itemize}
\item the analysis of group-level under- and overconfidence across different levels of self-reported expertise,
%\item confidence bands for the crowd belief,
\item accurate probability forecasts, and
\item many question-specific quantities that have easy interpretations and can be used to gain novel insight in the social sciences. 
\end{itemize}

@@

%Two commonly desired attributes of an aggregator are \textit{calibration} and \textit{sharpness}:

%Sharpness, on the other hand, is a statement about the spread of the forecasts. The aggregator is said to be sharp if the resulting forecasts are near 0 and 1. These two attributes are typically combined into a single goal of finding an aggregator that maximizes sharpness subject to calibration (see, e.g., \citet{gneiting2004verifying, raftery2005using}).

%
%
%
%
%
%\noindent
%\\
%Element 1: MAIN THEME: The broad theme is subjective probability aggregation. Use a catchy initial phrase: "Individuals can differ radically from one another in the precision to which they are able to  evaluate the likelihood of a future event happening." This is makes a statement about human cognitive process and does not include any odd terms. \\
%Element 2: WHY CARE: Explain the importance of probability aggregation (in a static context) for decision making. Explain why time-series advice is more realistic. (1 REABSTC-LogN) Lack of previous literature on time-series, subjective probability. (2 REASON)  \\
%Element 3: SUMMARY OF LITERATURE: See the guidelines. We want to to site work done on static case. There is also some work done on repeated forecasts, i.e. see raftery's paper.\\
%Element 4: MAIN CONTRIBUTION: "Our main contribution arises from our abilities ... " Advertise the data agin.\\
%Element 5: WE NEED A CLEAR INDICATION OF THE FOLLOWING:\\
%(a) The core research question: The main goal of this study is to find an interpretable model for detecting a a calibrated crowd belief.\\
%(b) More specific research objectives: With such a model we can make predictions, answer questions about bias-vs-expertise, and evaluate question difficulty. It may be a good idea to use bullet points here for this. \\
%(c) Context: Geopolitcal events. Describe the data we have briefly.\\
%(d) Units of analysis: Group of experts with different self-reported expertise; different questions in the dataset.\\
%Element 6: STRUCTURE OF THE PAPER\\
%%
%\ville{Much of the earlier paper can be included: define calibration and sharpness, normality around the calibrated crowd belief}
%
%\ville{Not studied in time-series - to our knowledge: REQUIRES LITERATURE REVIEW}
%
%\ville{Define calibration}
%
%\ville{Time series allows many more predictions and hence study of calibration -- this is not possible in static context where we would have only 166 predictions. Note that this method can still be applied to any given time point.}
%
%\ville{what are we hoping to learn from the data (applications) and why}
%
%\ville{Structure of the paper: theory, results (insight + precition)}
%
\section{Geopolitical Forecasting Data}
\label{data}
The data collection began with a recruitment of 2,365 forecasters ranging from graduate students to forecasting and political science faculty and practitioners. The recruiting was made from professional societies, research centers, alumni associations, science bloggers, and word of mouth. Requirements included at least a Bachelor's degree and completion of psychological and political tests that took roughly two hours. These measures assessed cognitive styles, cognitive abilities, personality traits, political attitudes, and real-world knowledge. The experts were asked to give probability forecasts (to the second decimal point) and to self-assess their level of expertise (measured on a 1-to-5 scale with 1 = Not At All Expert and 5 = Extremely Expert) on a number of 166 geopolitical binary events taking place between September 29, 2011 and May 8, 2013. Each question was active for a period of time during which the participating experts were allowed to update their forecasts as frequently as they liked without being penalized. The forecasters knew that their probability estimates would be assessed for accuracy using Brier scores\footnote{The Brier score is the squared distance between the probability forecast and the event indicator that equals 1.0 or 0.0 depending on whether the event happened or not, respectively. See \citet{Brier} for the original introduction.}. This incentivized them to report their true beliefs instead of attempting to game the system (\citet{winkler1968good}). In addition to receiving \$150 for meeting minimum participation requirements that did not depend on prediction accuracy, the forecasters also received status rewards for their performance via leader-boards displaying Brier scores for the top 20 forecasters. Since a typical expert participated only in a small subset of the 166 questions, the forecasters are considered indistinguishable conditional on the self-reported level of expertise.
Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data. Notice that the distribution of the self-reported expertise is skewed to the right
%: some questions involved very few experts who associated themselves with the highest level of expertise. Notice also 
and that some questions remained active longer than others.  For more details on the data and its collection see \citet{ungar2012good}.

To illustrate the nature of the data with some concrete examples, Figures \ref{Example1} and \ref{Example2} show scatterplots of the probability forecasts given for (a) \textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}, and (b) \textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}. The points have been jittered slightly to make overlaps visible.
%Overall, Figures \ref{Example1} and \ref{Example2} contain 872 forecasts made by 710 experts and 1,860 forecasts made by 1,376 experts, respectively. 
The darkness of the points are positively associated with the self-reported expertise levels. Given that the European bailout fund was ratified before 1 November 2011 and that the Nikkei 225 index finished trading at around 8,700 on 30 September 2011, the general trend of the probability forecasts converge to the correct answers. There is, however, much disagreement among the experts. This disagreement persists even near the closing dates of the questions. Unfortunately, the data is not only noisy but also quite sparse. As mentioned before, the forecasters were allowed to update their forecast as long as the questions remained active. Updating, however, was done on a very infrequent basis. An expert gave, on average, only 0.0159 forecasts per day. A typical expert, therefore, made a forecast about every 62 days. As a result the average response rate was around 13.5 forecasts per day from a large group of experts.

@@

\hspace{-0.5em}
\subfigure[\textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}]{
\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1}
%\hspace*{-1em}  \includegraphics[width= 1.5in]{CalibrationPLATTSPH10}
%\caption{$\beta = 1$, $\alpha = 0$}
\label{Example1}
} \hspace{0.5em}
\subfigure[\textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}]{
\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2}
%\caption{\textit{Sample-Then-Calibrate} with logarithmic score}
\label{Example2}
}

@@

%
% For instance, two of the questions were
%\begin{quote}\textit{
%Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011? %1021
%}\end{quote}
%and
%\begin{quote}\textit{
%Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011? %1011
%}
%\end{quote}

\section{Model}
\label{model}
Let $p_{i,t,k} \in (0,1)$ be the probability forecast given by the $i$th expert at time $t$ for the $k$th question, where $i = i, \dots, I_k$, $t = 1, \dots, T_k$, and $k = 1, \dots, K$. Denote the logit-probabilities with $Y_{i,t,k} = \logit(p_{i,t,k}) = \log \left( \frac{p_{i,t,k}}{1-p_{i,t,k}}\right) \in \R$ and collect the logit-probability forecasts given for question $k$ at time $t$ into a vector $\boldsymbol{Y}_{t,k} = [Y_{1,t,k}\ Y_{2,t,k}\ \dots\ Y_{I_{k},t,k}]^T$. Partition the experts into $J$ groups based on some individual feature, such as self-reported expertise, with each group sharing a common multiplicative bias term $b_{j} \in \R$ for $j = 1, \dots, J$. Collect these bias terms into a bias vector $\boldsymbol{b} = [b_{1}\ b_{2}\ \dots\ b_{J}]^T$ and let $\boldsymbol{M}_k$ be a $I_k \times J$ matrix denoting  the group-memberships of the forecasters in question $k$. That is, if the $i$th expert participating in the $k$th question belongs to the $j$th group, then the $i$th row of $\boldsymbol{M}_k$ is the $j$th standard basis vector, $\boldsymbol{e}_j$. The bias vector $\boldsymbol{b}$ does not include a subindex because it is considered shared among all the $K$ questions. To secure model identifiability, it is sufficient to share only one of the elements of $\boldsymbol{b}$ among the questions. This element defines a baseline under which it is possible to estimate the remaining $J-1$ bias terms separately within each of the questions. In this paper, however, the entire vector $\boldsymbol{b}$ is shared because some of the questions in our real-world data set involve very few experts with the highest level of self-reported expertise. Under this notation, the model for the $k$th question can be expressed as 
\begin{eqnarray}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} X_{t, k} + \boldsymbol{v}_{t, k} \label{observedpr} \\
X_{t, k} &=& \gamma_k X_{t-1, k} + w_{t, k} \label{hiddenpr}\\
X_{0,k} &\sim& \mathcal{N}(\mu_0, \sigma^2_0) \nonumber
\end{eqnarray}
where Equation (\ref{observedpr}) denotes the observed process, Equation (\ref{hiddenpr}) shows the hidden process that is driven by the constant $\gamma_k \in \R$, the error terms are independent and identically distributed normal random variables with mean zero
\begin{eqnarray*}
\boldsymbol{v}_{t, k} | \sigma^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}(0, \tau^2_k),
\end{eqnarray*}
and $(\mu_0, \sigma_0^2) \in (\R, \R^+)$ are hyper-parameters chosen a priori. The hidden state $X_{t,k}$ represents the sharp and calibrated logit-probability for the $k$th event given the information available up to and including time $t$. To make this more specific, let $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. If $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ is a filtration representing the information available up to and including a given time point, then $\E[Z_k | \mathcal{F}_{t,k}] = \P(Z_k = 1| \mathcal{F}_{t,k}) = \logit^{-1}(X_{t,k})$. These probabilities are made estimable by assuming that the observed process is associated with the filtration $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ and hence can be used as a proxy for the amount of information available at any given time point. Therefore it is not unreasonable to interpreted the hidden process as the crowd belief across time. 
%Each modeling task should begin with observations about the nature of the question. With a good understanding of the question, the modeler can proceed to construct a model that has an interpretable and intuitive connection to the real world.

@@

%
%\textit{Random noise, $\sigma^2_k$:} The variability of each observed value $Y_{i,t,k}$ around the biased hidden state $b_j \tilde{X}_{t,k}$. This can be thought of as being caused by momentary over-optimism (or pessimism), false beliefs, or other misconceptions. 
%Note that only the most informed expert making an objective estimation based solely on facts can be expected to give $X_t$ as his prediction. Unfortunately, human experts hardly make such objective estimates. Therefore this noise component is an important part of our model. 
%
%\textit{Random signal, $\tau^2_k$:} This variability is caused by changes in the amount of true information that is available to the experts. It allows the model to incorporate changes or shocks to the underlying circumstances that ultimately decide the outcome of the event. 
%
%\textit{Systematic noise, $b_{j}$:} This is a bias in the observed values given by the $j$th most expert group and can be interpreted as groupwise over- or under-confidence that pulls the observed values away from the hidden state values. 
%
%\textit{Systematic signal, $\gamma_k$:} Since the event associated with the $k$th question in our dataset realizes by time $T_k$, the uncertainty in the question reduces as the current time point $t$ approaches the final time point $T_k$. This reduction in uncertainty can be viewed as a constant signal stream that pushes the hidden process towards either extreme as more time passes by.

%These interpretations have been summarized in Table \ref{signalnoise}.
%\begin{table}
%\centering
%\begin{tabular}{| c | c |  c | }\hline
% Model Factors & Signal & Noise \\ 
%Random & $\tau^2_k$ & $\sigma^2_k$ \\ 
%Systematic & $\gamma_k$ & $b_{j,k}$ \\ \hline
%\end{tabular}
%\caption{The signal/noise structure of the dynamic linear model}
%\label{signalnoise}
%\end{table}

\section{Model Estimation}
\label{identifiability}
%Unfortunately, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable. To see this, let $a$ be any positive constant. Then $(\boldsymbol{b} a, X_{t,k}/a, a \tau) \neq (\boldsymbol{b}, X_{t,k}, \tau)$ yield the same likelihood. As a result, the parameter values tend to drift during the estimation process.  Several solutions are as follows.
%
%\begin{enumerate}
%\item[a.]  Place informative prior distributions on some of the non-identifiable parameters. These prior distributions, however, must be highly informative in order to provide identifiability. Since studying the expertise biases is one of the research objectives of this paper, placing a highly restrictive prior distribution on $\boldsymbol{b}$ is not satisfactory. Having strong apriori knowledge of $\tau_k$, on the other hand, seems difficult to justify.  
%\item[b.] The parameters could be constrained to reduce the number of degrees of freedom in the model. A popular choice is to fix one of the bias terms to some constant. For instance, the third expertise group could be chosen as the baseline group by fixing $b_3 = 1$. This, however, assumes that the third expertise group is unbiased. Such an assumption is hardly realistic and would hinder our analysis of under- and overconfidence among the expertise groups. 
%\item[c.] Reduce the degrees of freedom by introducing more information to the modeling process. 
% \end{enumerate}
% 
%The next two subsections introduce two hierarchical approaches that combine solutions b. and c. to restore identifiability without sacrificing the interpretability of the model. The first approach, called  \textit{Sample-Then-Calibrate}, is flexible and results into an efficient sampling algorithm. The second approach, named \textit{Sample-Only}, on the other hand, produces uncertainty estimates for all the unknown parameters in a coherent, fully Bayesian manner.

%

@@

%The second approach, named \textit{Sample-Only} (SO), on the other hand, produces uncertainty estimates for all the unknown parameters in a fully Bayesian manner.

@@

%In terms of the prior distributions, we assume that $\alpha$ and $\beta$ are a priori independent with 
%\begin{eqnarray*}
%\beta &\sim& \text{Cauchy}(1, 1/2.5)\\
%\alpha &\sim& \text{Cauchy}(0, 10),
%\end{eqnarray*}
%which are essentially the weakly informative default priors introduced in \citet{gelman2008weakly}. The only difference is in the centering of the prior $p(\beta)$. \citet{gelman2008weakly} recommend centering this prior at 0. Since a priori we have no reason to assume under- or overconfidence of the $\tilde{X}_{t,k}$ process, the prior $p(\beta)$ should be centered at 1 instead. 
%
%Unlike the \textit{Sample-Then-Calibrate} approach, the \textit{Sample-Only} approach models the uncertainty in estimating $(\beta, \alpha)$. The \textit{Sample-Only} approach also results into an algorithm that produces all estimates from a single procedure; hence, making it appear more coherent.  The \textit{Sample-Then-Calibrate} approach, on other hand, enjoys the flexibility to choose a score function, $S$, that suits the application at hand (see \citet{Gneiting04strictlyproper}). Furthermore, the  \textit{Sample-Then-Calibrate} approach results into a slightly more efficient sampling procedure.

%Even though both approaches assume a hierarchical model with a common $\boldsymbol{b}$ across the questions,

>>>>

\begin{aug}
\author{\fnms{Ville A.} \snm{Satop\"a\"a}\corref{}\ead[label=e1]{satopaa@wharton.upenn.edu}},
\author{\fnms{Shane T.} \snm{Jensen}\ead[label=e3]{stjensen@wharton.upenn.edu}},
\author{\fnms{Lyle H.} \snm{Ungar}\ead[label=e2]{ungar@cis.upenn.edu}},
\author{\fnms{Barbara A.} \snm{Mellers}\ead[label=e4]{mellers@wharton.upenn.edu}},
\and
\author{\fnms{Phil E.} \snm{Tetlock}\ead[label=e5]{tetlock@wharton.upenn.edu}}

@@

 \affiliation{Department of Statistics, The Wharton School of the University of Pennsylvania\\   \printead{e1,e3}}
 \affiliation{Department of Computer and Information Science, University of Pennsylvania\\ \printead{e2}}
 \affiliation{Department of Psychology, University of Pennsylvania\\ \printead{e4,e5}}
 \address{Philadelphia, PA 19104- 6340, USA\\  \printead{e1}}

@@

\section{Introduction}
\label{intro}
Individual experts can differ radically from one another in their abilities to assess probabilities of future events. The quality of their assessments is typically evaluated with an important attribute called \textit{calibration}. An expert is said to be well-calibrated if 
the proportion of events that actually occurred is, say, 60\% for all those events for which he assessed a probability of 0.60. Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), some personal differences among the experts have been found to associate with calibration. In particular, \citet{wright1994coherence} argue that a higher self-rated expertise on the subject matter implies better calibration.

Unfortunately, calibration by itself does not guarantee usefulness of a probability assessment. For instance, if the empirical frequency of  the last 10 years of a relatively stationary process is 45\%, the expert can assess a constant probability of 0.45 and turn out well-calibrated. His probability, however, is not very useful in decision making. In general, a probability assessments that is near 0.0 or 1.0, is considered more informative, and an expert who makes such assessments is called \textit{sharp}. In the forecasting literature sharpness is often combined with calibration into a single goal of maximizing sharpness subject to calibration (see, e.g., \citet{raftery2005using, Ranjan08}). An expert that meets this goal is considered to give perfect probability forecasts (\citet{gneiting2004verifying}). Even though expecting perfect forecasts from an expert is not reasonable in practice, this criterion has led to a wide range of novel and insightful observations in probability forecasting.

@@

Therefore many complex areas of probability aggregation still remain rather unexplored. For instance, consider an investor aiming to assess whether a stock index will finish trading at or above some threshold on a given date. In order to maximize his overall predictive accuracy, he may consult a group of experts every day and adjust his estimate of the aggregate probability accordingly. Since the experts are allowed to update their probability assessments on a daily basis, the aggregation should be performed by taking into account the correlation and time-series nature of their advice. Analyzing this kind of future events under daily updating of the forecasts is the main focus of this paper. The core research goal is to construct an interpretable time-series model that incorporates self-reported expertise in order to estimate a sharp and well-calibrated version of the crowd belief. The model is then used for
\begin{itemize}
\item the analysis of group-level under- and overconfidence across different levels of self-reported expertise,
%\item confidence bands for the crowd belief,
\item accurate probability forecasts, and
\item many question-specific quantities that have easy interpretations and can be used to gain novel insight in the social sciences. 
\end{itemize}

@@

\section{Geopolitical Forecasting Data}
\label{data}
The data collection began with a recruitment of 2,365 forecasters ranging from graduate students to forecasting and political science faculty and practitioners. The recruiting was made from professional societies, research centers, alumni associations, science bloggers, and word of mouth. Requirements included at least a Bachelor's degree and completion of psychological and political tests that took roughly two hours. These measures assessed cognitive styles, cognitive abilities, personality traits, political attitudes, and real-world knowledge. The experts were asked to give probability forecasts (to the second decimal point) and to self-assess their level of expertise (measured on a 1-to-5 scale with 1 = Not At All Expert and 5 = Extremely Expert) on a number of 166 geopolitical binary events taking place between September 29, 2011 and May 8, 2013. Each question was active for a period of time during which the participating experts were allowed to update their forecasts as frequently as they liked without being penalized. The forecasters knew that their probability estimates would be assessed for accuracy using Brier scores\footnote{The Brier score is the squared distance between the probability forecast and the event indicator that equals 1.0 or 0.0 depending on whether the event happened or not, respectively. See \citet{Brier} for the original introduction.}. This incentivized them to report their true beliefs instead of attempting to game the system (\citet{winkler1968good}). In addition to receiving \$150 for meeting minimum participation requirements that did not depend on prediction accuracy, the forecasters also received status rewards for their performance via leader-boards displaying Brier scores for the top 20 forecasters. Since a typical expert participated only in a small subset of the 166 questions, the forecasters are considered indistinguishable conditional on the self-reported level of expertise.
Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data. Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.  For more details on the data and its collection see \citet{ungar2012good}.

To illustrate the nature of the data with some concrete examples, Figures \ref{Example1} and \ref{Example2} show scatterplots of the probability forecasts given for (a) \textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}, and (b) \textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}. The points have been jittered slightly to make overlaps visible. The darkness of the points are positively associated with the self-reported expertise levels. Given that the European bailout fund was ratified before 1 November 2011 and that the Nikkei 225 index finished trading at around 8,700 on 30 September 2011, the general trend of the probability forecasts converge to the correct answers. There is, however, much disagreement among the experts. This disagreement persists even near the closing dates of the questions. Unfortunately, the data is not only noisy but also quite sparse. As mentioned before, the forecasters were allowed to update their forecast as long as the questions remained active. Updating, however, was done on a very infrequent basis. An expert gave, on average, only 0.0159 forecasts per day. A typical expert, therefore, made a forecast about every 62 days. As a result the average response rate was around 13.5 forecasts per day from a large group of experts.

@@

\hspace{-0.5em}
\subfigure[\textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}]{
\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot1}
\label{Example1}
} \hspace{0.5em}
\subfigure[\textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}]{
\hspace*{-1em}  \includegraphics[width= 0.49\textwidth]{Figures/ExamplePlot2}
\label{Example2}
}

@@

\section{Model}
\label{model}
Let $p_{i,t,k} \in (0,1)$ be the probability forecast given by the $i$th expert at time $t$ for the $k$th question, where $i = i, \dots, I_k$, $t = 1, \dots, T_k$, and $k = 1, \dots, K$. Denote the logit-probabilities with $Y_{i,t,k} = \logit(p_{i,t,k}) = \log \left( \frac{p_{i,t,k}}{1-p_{i,t,k}}\right) \in \R$ and collect the logit-probability forecasts given for question $k$ at time $t$ into a vector $\boldsymbol{Y}_{t,k} = [Y_{1,t,k}\ Y_{2,t,k}\ \dots\ Y_{I_{k},t,k}]^T$. Partition the experts into $J$ groups based on some individual feature, such as self-reported expertise, with each group sharing a common multiplicative bias term $b_{j} \in \R$ for $j = 1, \dots, J$. Collect these bias terms into a bias vector $\boldsymbol{b} = [b_{1}\ b_{2}\ \dots\ b_{J}]^T$ and let $\boldsymbol{M}_k$ be a $I_k \times J$ matrix denoting  the group-memberships of the forecasters in question $k$. That is, if the $i$th expert participating in the $k$th question belongs to the $j$th group, then the $i$th row of $\boldsymbol{M}_k$ is the $j$th standard basis vector, $\boldsymbol{e}_j$. The bias vector $\boldsymbol{b}$ does not include a subindex because it is considered shared among all the $K$ questions. To secure model identifiability, it is sufficient to share only one of the elements of $\boldsymbol{b}$ among the questions. This element defines a baseline under which it is possible to estimate the remaining $J-1$ bias terms separately within each of the questions. In this paper, however, the entire vector $\boldsymbol{b}$ is shared because some of the questions in our real-world data set involve very few experts with the highest level of self-reported expertise. Under this notation, the model for the $k$th question can be expressed as 
\begin{eqnarray}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} X_{t, k} + \boldsymbol{v}_{t, k} \label{observedpr} \\
X_{t, k} &=& \gamma_k X_{t-1, k} + w_{t, k} \label{hiddenpr}\\
X_{0,k} &\sim& \mathcal{N}(\mu_0, \sigma^2_0) \nonumber
\end{eqnarray}
where Equation (\ref{observedpr}) denotes the observed process, Equation (\ref{hiddenpr}) shows the hidden process that is driven by the constant $\gamma_k \in \R$, the error terms are independent and identically distributed normal random variables with mean zero
\begin{eqnarray*}
\boldsymbol{v}_{t, k} | \sigma^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}(0, \tau^2_k),
\end{eqnarray*}
and $(\mu_0, \sigma_0^2) \in (\R, \R^+)$ are hyper-parameters chosen a priori. The hidden state $X_{t,k}$ represents the sharp and calibrated logit-probability for the $k$th event given the information available up to and including time $t$. To make this more specific, let $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. If $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ is a filtration representing the information available up to and including a given time point, then $\E[Z_k | \mathcal{F}_{t,k}] = \P(Z_k = 1| \mathcal{F}_{t,k}) = \logit^{-1}(X_{t,k})$. These probabilities are made estimable by assuming that the observed process is associated with the filtration $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ and hence can be used as a proxy for the amount of information available at any given time point. Therefore it is not unreasonable to interpreted the hidden process as the crowd belief across time.

@@

\section{Model Estimation}
\label{identifiability}

@@

@@

\section{Extension: Polychotomous Outcomes}
If the future event can take upon $M > 2$ possible outcomes, the hidden state $X_{t,k}$ must be extended to a vector of size $M-1$. One of the outcomes, e.g., the $M$th one, is chosen as the base-case.  This gives us a total of $M-1$ observed processes and one hidden process
\begin{eqnarray*}
\boldsymbol{Y}_{1,t,k} &=&  \boldsymbol{M}_k\boldsymbol{b}_{1}X_{1,t,k} + \boldsymbol{v}_{1,t,k}\\
\boldsymbol{Y}_{2,t,k} &=&  \boldsymbol{M}_k\boldsymbol{b}_{2}X_{2,t,k} + \boldsymbol{v}_{2,t,k}\\
&\vdots&\\
\boldsymbol{Y}_{M-1,t,k} &=&  \boldsymbol{M}_k\boldsymbol{b}_{M-1}X_{M-1,t,k} + \boldsymbol{v}_{M-1,t, k}\\
\boldsymbol{X}_{t,k} &=& \boldsymbol{\gamma}^T_k \boldsymbol{I}_{M-1} \boldsymbol{X}_{t-1,k} + \boldsymbol{w}_{t,k}
\end{eqnarray*}
where $\boldsymbol{X}_{t,k}$ is a $(M-1) \times 1$ matrix of calibrated logit-probabilities. Notice that each process has a separate bias vector. The error terms are independent and identically distributed multivariate normal random variables 
\begin{eqnarray*}
\boldsymbol{v}_{m, t, k} | \sigma^2_{j,k} &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k} \left(\boldsymbol{0}, \sigma^2_{j,k} \boldsymbol{I}_{I_k} \right)\\
\boldsymbol{w}_{t,k} | \tau_k^2 &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{M-1} \left(\boldsymbol{0}, \tau_k^2 \boldsymbol{I}_{M-1} \right)
\end{eqnarray*}
On the left-hand side of observed process
\begin{eqnarray*}
\boldsymbol{Y}_{m,t,k} &=&  \left[ \begin{matrix} Y_{m,1,t,k} & Y_{m,2,t,k} & \dots & Y_{m,I_k,t,k} \end{matrix} \right]^T\\
&=&  \left[ \begin{matrix} \log \left( \frac{p_{m,1,t,k}}{p_{M,1,t,k}} \right) &  \log \left( \frac{p_{m,2,t,k}}{p_{M,2,t,k}} \right) & \dots &  \log \left( \frac{p_{m,I_k,t}}{p_{M,I_k,t,k}} \right) \end{matrix} \right]^T
\end{eqnarray*}
is a $I_k \times 1$ matrix of the expert logit-probabilities for the $m$th outcome of the $k$th question at time $t$, and on the left-hand side of the hidden process
\begin{eqnarray*}
\boldsymbol{X}_{t,k} &=&  \left[ \begin{matrix}X_{1,t,k} &X_{2,t,k} & \dots &X_{M-1,t,k} \end{matrix} \right]^T
\end{eqnarray*}
is the matrix of sharp and calibrated logit-probabilities at time $t$. Choosing one of the outcomes as the base-case, ensures that the probabilities will sum to one at any given time point. Since this multinomial extension is equivalent to having $M-1$ independent binary-outcome models (see Section \ref{model}), the estimation can be done separately for each outcome. Therefore, even though this paper focuses on modeling binary events, all the properties and discussion generalize trivially to the multi-outcome case.
